Parsing Tweets into Universal Dependencies

### 2017-12-11 paper.tex
Noah

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets. Using the extended guidelines \yicomment{including tokenization, POS tagging and parsing? (need to mention the annotation process/pipeline system not only consists parsing in abstract)}, we create
a new tweet treebank ({\sc Tweebank v2}), which has 3,550 tweets\yicomment{add tweet \#}\yjcomment{didn't see necessity of mentioning \# of tweets} and 55,607 tokens with
labeled attachments. 
It is more than four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreement
among our annotations and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over \nascomment{which
  comparison is this?}\yjcomment{my bad, it was fixed.} the baseline
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank \citep[PTB]{Marcus93buildinga}.
Stanford dependencies were converted afterwards in their work  \nascomment{by who?}\yjcomment{by themselves}. 
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Universal dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines. Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) universal dependencies.
We characterize the disagreement among our annotations \yicomment{annotation(s) and disagreement(s). check abstract, same sentence, different singularity/plurality. I think should be `disagreement' and `annotations'} and find that
consistent annotation \yicomment{annotations?} is still challenging to deliver even with
the extended guidelines.

@@

Based on these annotations, we designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.6 points \nascomment{again, not sure what this is quoting}\yjcomment{fixed} in LAS over the
baseline parser can be achieved with our distillation method and it outperforms
other state-of-the-art parsers in both accuracy and running speed.

@@

\item We introduce and evaluate a \yicomment{pipeline} system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  Experimental results show
it performs better than a stack of the state-of-the-art alternatives.
\nascomment{say it
  performs better than existing systems?}\yjcomment{added.}

@@

We first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
which is \yicomment{was?} the largest \yicomment{unlabelled?} Twitter dependencies annotation
(\S\ref{sec:tweebank}).
Then we introduce the differences in our tokenization
(\S\ref{sec:tok-anno}) and part-of-speech (\S\ref{sec:pos-anno}) (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011}, respectively, on which {\sc Tweebank v1} was built. 
We describe our effort of adapting the
UD conventions to cover tweet-specific constructions (\S\ref{sec:ud-tweet}). 
Finally, we present our process of creating a new tweet treebank, {\sc
  Tweebank v2}, and characterize
the difficulties in reaching consistent annotations (\S\ref{sec:anno-process}).

@@

The annotation effort we describe stands in contrast to \yicomment{the?} previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
12,149 tokens.  Their annotations added \yicomment{unlabelled?} dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after rule-based tokenization \citep{ICWSM101540}.
Kong et al.~also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

Kong et al.'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out mostly by nonnative speakers \yicomment{should not emphasize non-native speakers (v2 also built by non-native speakers), but emphasize annotators with only cursory training, no clear annotation guidelines, no consensus/agreement on controversial cases, allow annotators to underspecify part of the tree and allow multiple different trees for the same tweets}, in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic''.
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

\subsection{Tokenization}\label{sec:tok-anno}
Our tokenization strategy lies between the strategy of
\citet{ICWSM101540} and that of UD.
There is a significant \yicomment{significant?} tradeoff between preservation of original tweet content and respecting
the UD guidelines.

The regex-based tokenizer of \citet{ICWSM101540}---which was 
originally designed for an exploratory search interface called
TweetMotif, not for NLP---preserves most whitespace-delimited tokens, including 
hashtags, at-mentions, emoticons, and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD
tokenization,\footnote{\url{http://universaldependencies.org/u/overview/tokenization.html}}
in order to better serve dependency annotation, treats each syntactic word as a token.
They therefore more aggressively split
clitics from contractions (e.g., {\it  gonna} is tokenized as {\it gon} and {\it na}; {\it its}
is tokenized as {\it it} and {\it s} when {\it s} is a copula)\yicomment{are words like gonna called contractions, or do they have another term?}. But acronyms are not touched
in the UD tokenization guidelines. Thus, we follow the UD tokenization for contractions
and left acronyms like {\em idc} as a single token. \nascomment{unless we make a change, I think we need to say
  here that acronyms like \emph{idc} are left as a single token}
\yjcomment{mentioned acronymns, we don't have the problem of normalization, cause
normalization was just annotated for data study.}

@@

Before turning to UD annotations, we (re)annotated the data with 
POS \yicomment{tags?}, for consistency with other UD efforts,
which adopt the universal POS tagset of \newcite{PETROV12.274}.
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases, conflicts arose between the UD English 
treebank conventions \cite[][UD\_English]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT}.  \yicomment{The conflicts should refer to a {\bf noncorresponding tag} from that of gimpel to UDEnglish, cuz they dont have the same tagset.} In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

\paragraph{Acronym abbreviations.}  How should we syntactically
analyze acronym tokens like \emph{idc}
(abbreviating ``I don't care'')
 and \emph{rn} (``right now'')?  Should
they be decomposed into their component words, and if so should those
words be ``normalized'' into explicitly spelled out intermediate
forms? \yicomment{this sentence should be deleted or moved to tokenization?} \yicomment{As we already know from previous sections, \emph{idc} is not tokenized and tagged as verb, we follow ...} \yicomment{Normalization is discussed in the 2.5.2 Annotation Process. Actually we did do normalization, but just as auxiliary info, didn't use it for any model.}\yjcomment{i think acronym is a different story to tokenization.}
We follow \citet{kong-EtAl:2014:EMNLP2014} and annotate their
syntax as a single word without normalization. Their syntactic functions
are decided according to their context. \citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and argued that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated step.\yicomment{Again we did normalization!}
% \nascomment{maybe we owe citations to
%   \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
%   better papers to cite}

@@

The major characteristic that distinguishes tweets from standard texts is that
a large proportion of tokens don't carry any syntactic function.
In our annotation, there are five types of non-syntactic tokens: 
sentiment emoticons, retweet markers and their following at-mentions, topical hashtags, referential URLs, and
truncated words.\footnote{The tweets we analyze have at most 140
  characters. Although Twitter has doubled the tweet length limit to
  280 characters since our analysis, we believe this type of token
  will still remain.}  Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.55\% of all tokens are non-syntactic; detailed statistics can be found in Table \ref{tbl:non-synt-prop}.
\yjcomment{I've merged retweet marker and retweet at-mention into one.}

It is important to note that these types may, in some
contexts, have syntactic functions.
For example, besides being a discourse marker, \emph{RT} can
abbreviate the verb {\tt retweet}, and emoticons and hashtags may be
used as content words within a sentence\yicomment{at-mentions can be a normal vocative proper noun}; see Figure
\ref{fig:informal-toks}.  Therefore, the criteria for annotating a
token as non-syntactic must be context-dependent.
%To our knowledge, this issue has not previously been addressed in UD
%annotation efforts \nascomment{check!}\yjcomment{it's not safe to claim that.}.

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:\yicomment{shall we mention how we segment sentences? Seems relevant here}\yjcomment{i don't want to touch sentence segmentation, it will makes the paper more messy.}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependencies annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms \yicomment{and preserve the original tweets as much as possible},
we include non-syntactic tokens in our annotation following the
conventions above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear %\yicomment{much clearer}
definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}\yjcomment{i wouldn't like the readers to think in this direction.}.  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Constructions handled by UD.}  A number of constructions
that are especially common in tweets are well handled by UD
conventions: ellipsis, irregular word orders, and paratactic sentences \yicomment{or phrases/components}
not explicitly delineated by punctuation.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is an at-mention %(followed by a colon \nascomment{?})
as a sign that a tweet is a reply to a tweet by the mentioned user\yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}\yjcomment{i didn't see any difference between reply and mention, they all starts with a @username.}.  We
treat these at-mentions as vocative expressions, labeling them as
\emph{proper noun} (\texttt{PROPN})\yicomment{here is confusing that PROPN could be thought as a label in dependencies.} and attaching them to the
following predicate with the label \emph{vocative} (see Figure \ref{fig:informal-toks} for an example) \yicomment{just as UD guidelines}.

@@

\subsection{\sc Tweebank v2}\label{sec:anno-process}
Following the guidelines presented above, we create a new annotated\yicomment{del?} Twitter \yicomment{Twitter dependency?} treebank, which we call {\sc Tweebank v2}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is few disagreement on the tokenization annotation.
The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) \yicomment{noun phrases or proper noun phrases, do we need example here?}\yjcomment{i wasn't counting PROPN and NOUN confusion, but any type of confusion related with PROPN} and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' \yicomment{need to relate to previous error analysis, entity names maybe?}can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. \yicomment{need to say the reason why dependency inter-annotator agreement is low? 1. POS disagreement lead to big difference in understanding the semantics. 2. Some special constructions like ellipsis lead to different view of tweets. Fig3 is an example of 2.}An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.

>>>>

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets that affect tokenization,
part-of-speech tagging, and dependencies. Using the extended guidelines, we create
a new tweet treebank ({\sc Tweebank v2}) that is four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreements between our annotators
and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Nonetheless, using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over the un-ensembled baseline 
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank
\citep[PTB;][]{Marcus93buildinga}, enabling conversion to Stanford dependencies.
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Universal dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines. Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) universal dependencies.
We characterize the disagreements among our annotators and find that
consistent annotation is still challenging to deliver even with
the extended guidelines.

@@

Based on these annotations, we nonetheless designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.6 points in LAS over the
baseline parser can be achieved with our distillation method and it outperforms
other state-of-the-art parsers in both accuracy and speed.

@@

\item We introduce and evaluate a pipeline system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  Experimental results show
it performs better than a pipeline of the state-of-the-art alternatives.

@@

We first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
the largest Twitter dependency annotation effort to-date
(\S\ref{sec:tweebank}).
Then we introduce the differences in our tokenization
(\S\ref{sec:tok-anno}) and part-of-speech (\S\ref{sec:pos-anno}) (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011}, respectively, on which {\sc Tweebank v1} was built. 
We describe our effort of adapting the
UD conventions to cover tweet-specific constructions (\S\ref{sec:ud-tweet}). 
Finally, we present our process of creating a new tweet treebank, {\sc
  Tweebank v2}, and characterize
the difficulties in reaching consistent annotations (\S\ref{sec:anno-process}).

@@

The annotation effort we describe stands in contrast to the previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
12,149 tokens.  Their annotations added unlabeled dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after rule-based tokenization \citep{ICWSM101540}.
Kong et al.~also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

Kong et al.'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out by annotators with only cursory training, no clear
annotation guidelines, and no effort to achieve consensus on controversial
cases. Annotators were allowed to underspecify their analyses.
Most of the work was done in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic.''
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

\subsection{Tokenization}\label{sec:tok-anno}
Our tokenization strategy lies between the strategy of
\citet{ICWSM101540} and that of UD.
There is a tradeoff between preservation of original tweet content and respecting
the UD guidelines.

The regex-based tokenizer of \citet{ICWSM101540}---which was 
originally designed for an exploratory search interface called
TweetMotif, not for NLP---preserves most whitespace-delimited tokens, including 
hashtags, at-mentions, emoticons, and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD
tokenization,\footnote{\url{http://universaldependencies.org/u/overview/tokenization.html}}
in order to better serve dependency annotation, treats each syntactic word as a token.
They therefore more aggressively split
clitics from contractions (e.g., {\it  gonna} is tokenized as {\it gon} and {\it na}; {\it its}
is tokenized as {\it it} and {\it s} when {\it s} is a copula).
%\yicomment{are words like gonna called contractions, or do they have
%another term?}. \nascomment{yes}
But acronyms are not touched
in the UD tokenization guidelines. Thus, we follow the UD tokenization for contractions
and leave acronyms like {\em idc} as a single token. 
%\nascomment{unless we make a change, I think we need to say
%  here that acronyms like \emph{idc} are left as a single token}
%\yjcomment{mentioned acronymns, we don't have the problem of normalization, cause
% normalization was just annotated for data study.}

@@

Before turning to UD annotations, we (re)annotated the data with 
POS tags, for consistency with other UD efforts,
which adopt the universal POS tagset of \newcite{PETROV12.274}.
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases,  non-corresponding tag conflicts arose between the UD English 
treebank conventions \cite[UD\_English;][]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT}.  %\yicomment{The conflicts
                                %should refer to a {\bf
                                %noncorresponding tag} from that of
                                %gimpel to UDEnglish, cuz they dont
                                %have the same tagset.} 
In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

\paragraph{Acronym abbreviations.}  % How should we syntactically
% analyze acronym tokens like \emph{idc}
% (abbreviating ``I don't care'')
%  and \emph{rn} (``right now'')?  Should
% they be decomposed into their component words, and if so should those
% words be ``normalized'' into explicitly spelled out intermediate
% forms? \yicomment{this sentence should be deleted or moved to tokenization?} \yicomment{As we already know from previous sections, \emph{idc} is not tokenized and tagged as verb, we follow ...} \yicomment{Normalization is discussed in the 2.5.2 Annotation Process. Actually we did do normalization, but just as auxiliary info, didn't use it for any model.}\yjcomment{i think acronym is a different story to tokenization.}
\nascomment{removed redundant stuff here, please check}
We follow \citet{kong-EtAl:2014:EMNLP2014} and annotate the
syntax of an acronym as a single word without normalization. Their syntactic functions
are decided according to their context. \citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and argued that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated
step.\yicomment{Again we did normalization!} \nascomment{maybe just
  add ``for acronyms'' to the end of that last sentence?}
% \nascomment{maybe we owe citations to
%   \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
%   better papers to cite}

@@

The major characteristic that distinguishes tweets from standard texts is that
a large proportion of tokens don't carry any syntactic function.
In our annotation, there are five types of non-syntactic tokens: 
sentiment emoticons, retweet markers and their following at-mentions, topical hashtags, referential URLs, and
truncated words.\footnote{The tweets we analyze have at most 140
  characters. Although Twitter has doubled the tweet length limit to
  280 characters since our analysis, we believe this type of token
  will still remain.}  Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.55\% of all tokens are non-syntactic; detailed statistics can be found in Table \ref{tbl:non-synt-prop}.
%\yjcomment{I've merged retweet marker and retweet at-mention into one.}

It is important to note that these types may, in some
contexts, have syntactic functions.
For example, besides being a discourse marker, \emph{RT} can
abbreviate the verb {\tt retweet}, and emoticons and hashtags may be
used as content words within a sentence, and at-mentions can be normal vocative proper nouns; see Figure
\ref{fig:informal-toks}.  Therefore, the criteria for annotating a
token as non-syntactic must be context-dependent.
%To our knowledge, this issue has not previously been addressed in UD
%annotation efforts \nascomment{check!}\yjcomment{it's not safe to claim that.}.

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:
%\yicomment{shall we mention how we segment sentences? Seems relevant
%here}\yjcomment{i don't want to touch sentence segmentation, it will
%makes the paper more messy.}\nascomment{agree with Yijia}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependencies annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms and preserve the original tweets as much as possible,
we include non-syntactic tokens in our annotation following the
conventions above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear %\yicomment{much clearer}
definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object.
% \yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}
%\yjcomment{i wouldn't like the readers to think in this direction.}.
  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Constructions handled by UD.}  A number of constructions
that are especially common in tweets are well handled by UD
conventions: ellipsis, irregular word orders, and paratactic phrases
and sentences 
not explicitly delineated by punctuation.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is a vocative at-mention (sometimes a
signal that a tweet is a reply to a tweet by the mentioned user). % \yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}\yjcomment{i didn't see any difference between reply and mention, they all starts with a @username.}.
We
treat these at-mentions as vocative expressions, labeling them with
POS tag
\emph{proper noun} (\texttt{PROPN}) and attaching them to the
following predicate with the label \emph{vocative} as in UD guidelines (see Figure \ref{fig:informal-toks} for an example).

@@

\subsection{\sc Tweebank v2}\label{sec:anno-process}
Following the guidelines presented above, we create a new Twitter
dependency treebank, which we call {\sc Tweebank v2}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is very little disagreement on the
tokenization annotation \nascomment{still better to quantify}.
The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) %\yicomment{noun phrases or proper noun
                              %phrases, do we need example
                              %here?}\yjcomment{i wasn't counting
                              %PROPN and NOUN confusion, but any type
                              %of confusion related with PROPN}
                              %\nascomment{probably don't need examples}
and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' 
%\yicomment{need to relate to previous error analysis, entity names
%maybe?}
can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. 
%\yicomment{need to say the reason why dependency inter-annotator
%agreement is low? 1. POS disagreement lead to big difference in
%understanding the semantics. 2. Some special constructions like
%ellipsis lead to different view of tweets. Fig3 is an example of 2.}
%\nascomment{I don't think we have room to go deeper}
An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.
